Manuscript Figures

### 2021-03-18 src/manuscript_figures.Rmd
add supplementary figure about subtypes dependency on outcome

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
# todo: pass this as argument
settings <- yaml::read_yaml('/data/local/buyar/arcas/pancancer_multiomics_manuscript/settings.yaml')
source(settings$utility_script)
library(data.table)
library(ggplot2)
library(pbapply)
library(ggrepel)
library(ggpubr)
library(cowplot)
library(ggridges)
theme_set(ggpubr::theme_pubclean())
geneSetScoresFolder <- '/data/local/buyar/arcas/subtyping_paper/data/geneset_scores/singscore'
FigureFolder <- '/data/bimsbstatic/public/akalin/buyar/manuscript_figures_arcas'

@@

supFig_count <- supFig_count + 1
ggsave(paste0('SuppFigure', supFig_count, '.pancancer_subtype_prediction.pdf'), 
       plot = Fig, units = 'in', width = 18, height = 12)
SuppFigures[[supFig_count]] <- Fig
```

@@

## TCGA subtype informed clusters 
## Use-case: pan-gastrointestinal cancers
require(survminer)
require(cluster)
pr <- 'pan-gastrointestinal' 
experiment <- paste0(pr, '.cnv_gex_meth_mut')
M <- LFs$maui_factors[[experiment]]
st <- tcga_subtypes[match(rownames(M), bcr_patient_barcode)]$Subtype_Selected
# find subtype specific LFs 
dt <- get_factor_specific_variables(M, st)

@@

```{r print_figures, results='asis', include=TRUE, fig.width=10, fig.height=11}
base_url <- 'https://bimsbstatic.mdc-berlin.de'
for(i in 1:length(Figures)) {
  fig.url <- gsub("/data/bimsbstatic/public", base_url, file.path(FigureFolder, names(Figures)[i]))
  cat("# Figure",i,"[high-res image](",fig.url,")\n\n")
  print(Figures[[i]])
  cat("\n\n Figure",i,":",Legends[[i]],"\n\n")
}
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
# todo: pass this as argument
settings <- yaml::read_yaml('/data/local/buyar/arcas/pancancer_multiomics_manuscript/settings.yaml')
source(settings$utility_script)
library(data.table)
library(ggplot2)
library(pbapply)
library(ggrepel)
library(ggpubr)
library(cowplot)
library(ggridges)
library(ggcorrplot)
library(ComplexHeatmap)
library(pheatmap)
theme_set(ggpubr::theme_pubclean())
geneSetScoresFolder <- '/data/local/buyar/arcas/subtyping_paper/data/geneset_scores/singscore'
FigureFolder <- '/data/bimsbstatic/public/akalin/buyar/manuscript_figures_arcas'

@@

supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- Fig
fig.name <- paste0('SuppFigure', fig_count, '.pancancer_subtype_prediction.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = Fig, units = 'in', width = 18, height = 12)
SuppLegends[[fig_count]] <- "Predictive power of latent factors in distinguishing known TCGA cancer subtypes A-B) comparing MAUI with PCA/MOFA/MCIA. AUC: Area under the ROC curve computed on the test dataset (30% of held-out data) trained on 70% of the dataset using a Random Forest model. Each point represents one cancer subtype (pairwise one-sided Wilcoxon test on AUCs, alternative = maui AUC is not greater than others) C) Top maui LFs detected per cancer subtype (only displaying subtypes with at least 20 samples). Size/color/opacity of the points represent the average neural activation value of the LF for the corresponding subtype samples. D) Percentage of top feature types contributing to top predictive maui latent factor per cancer subtype."
```

@@

## TCGA subtype informed clusters 
## Use-case: pan-gastrointestinal cancers
require(survminer)
require(cluster)
pr <- 'pan-gastrointestinal' 
experiment <- paste0(pr, '.cnv_gex_meth_mut')
M <- remove_redundant_variables(LFs$maui_factors[[experiment]], perc = 99)
st <- tcga_subtypes[match(rownames(M), bcr_patient_barcode)]$Subtype_Selected
# find subtype specific LFs 
dt <- get_factor_specific_variables(M, st)

@@

```{r}
# Show subtypes depend on the objective (dependent variable) - they are dynamic, not universal or carved in stone 
# for a given cancer. There is not just one set of subtypes for a given cancer. Subtypes need to be phrased 
# along with the dependent variable used to come up with the subtypes. 
# Patients clustering depend on the objective of stratification
clin_factors <- c('age_at_initial_pathologic_diagnosis', 'gender', 'ajcc_pathologic_tumor_stage')
project <- 'pan-gastrointestinal'
datatype <- 'cnv_gex_meth_mut'
experiment <- paste0(project, '.', datatype)

M <- LFs$maui_factors[[experiment]]
M <- remove_redundant_variables(M, perc = 99)

clin <- process_clinical_covariates(rownames(M))
clin <- subset(clin, select = intersect(colnames(clin), clin_factors))

pfi_vars <- get_prognostic_covariates(surv, clin, M, event_col = 'PFI', duration_col = 'PFI.time')
pfi_clust <- get_kmeans_subtypes(M[,unique(pfi_vars$variable)], k_vals = 3:8, min_cluster_size = 0)
pfi_k <- names(which.min(sapply(colnames(pfi_clust), function(x) plot_survival(surv, pfi_clust[,x,drop = F], 'PFI.time', 'PFI', T))))

os_vars <- get_prognostic_covariates(surv, clin, M, event_col = 'OS', duration_col = 'OS.time')
os_clust <- get_kmeans_subtypes(M[,unique(os_vars$variable)], k_vals = 3:8, min_cluster_size = 0)
os_k <- names(which.min(sapply(colnames(os_clust), function(x) plot_survival(surv, os_clust[,x,drop = F], 
                                                                             'OS.time', 'OS', T))))

# combine with known pan-gi subtype annotation
df <- merge(pfi_clust[,pfi_k,drop = F], os_clust[,os_k,drop = F], by = 'row.names')
df$subtype <- tcga_subtypes[match(df$Row.names, bcr_patient_barcode)]$Subtype_Selected
df[is.na(df$subtype),]$subtype <- 'Unavailable'
df$msi <- MSI[match(df$Row.names, bcr_patient_barcode)]$MSI
df[is.na(df$msi),]$msi <- 'Unavailable'
rownames(df) <- df$Row.names
df$Row.names <- NULL
colnames(df) <- c('Overall Survival', 'Progression Free Survival', 'Molecular Subtype', 'MSI Status')
df <- data.frame(apply(df, 2, as.factor), check.names = F)
df <- df[order(df[,1]),]
m <- apply(df, 2, function(x) as.numeric(as.factor(x)))
rownames(m) <- rownames(df)

colors <- structure(1:7, names = c("1", "2", "3", "4", "5", "6", "7"))
p1 <- Heatmap(t(m), name = "cluster", col = colors, cluster_rows = F, cluster_columns = F,
              show_column_names = F, row_names_side = 'left',
              column_title = 'Pan-gastrointestinal Cancer Samples\n(ordered by Overall Survival Clusters)')
p1 <- grid::grid.grabExpr(ComplexHeatmap::draw(p1))
# compute pairwise AMI 
ami <- sapply(colnames(df), function(x) {
  sapply(colnames(df), function(y) {
    df <- df[!is.na(df[,x]) & !is.na(df[,y]),]
    aricode::AMI(df[,x], df[,y])
  })
})

p2 <- ggcorrplot::ggcorrplot(ami, legend.title = 'AMI', lab = T, type = 'lower', outline.color = 'white') +
  scale_fill_gradient(limits = c(0,1), low = 'white', high = 'red') + 
  guides(fill=guide_legend(title='AMI'))

Fig <- cowplot::plot_grid(p1, p2, labels = 'AUTO', nrow = 2)

supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- Fig
fig.name <- paste0('SuppFigure', fig_count, '.subtype_dependency.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = Fig, units = 'in', width = 18, height = 12)
SuppLegends[[fig_count]] <- "Demonstration of how cancer subtypes are context/objective dependent.
A) Heatmap of patient cluster memberships based on different objectives: optimum stratification based on overall survival or progression free survival, molecular homogeneity (TCGA pan-GI subtypes), or micro-satellite instability (MSI) status. Columns are ordered first by membership in the optimum Overall Survival cluster membership. B) Pairwise Adjusted Mutual Information (AMI) of cluster labels obtained via different patient stratification methods. The AMI value ranges from 0 to 1 and the higher the value, the more overlap between patient partitions."
```

```{r print_figures, results='asis', include=TRUE, fig.width=10, fig.height=11}
base_url <- 'https://bimsbstatic.mdc-berlin.de'
for(i in 1:length(Figures)) {
  fig.url <- gsub("/data/bimsbstatic/public", base_url, file.path(FigureFolder, names(Figures)[i]))
  cat("# Figure",i,"[high-res image](",fig.url,")\n\n")
  print(Figures[[i]])
  cat("\n\n Figure",i,":",Legends[[i]],"\n\n")
}

for(i in 1:length(SuppFigures)) {
  fig.url <- gsub("/data/bimsbstatic/public", base_url, file.path(FigureFolder, names(SuppFigures)[i]))
  cat("# Supplementary Figure",i,"[high-res image](",fig.url,")\n\n")
  print(SuppFigures[[i]])
  cat("\n\n Supplementary Figure",i,":",SuppLegends[[i]],"\n\n")
}
```
